Very Large Integers and polynomial inner products

### 2013-01-16 doc/paper_vli/vli.tex
paper

 THE LIB DOES
\begin{itemize}
\item The Very Large Integer library realizes large arithmetic , polynomials  and vector operation.  There are three 3 different modules interchangeable. 
\end{itemize}

\begin{enumerate}
\item  \begin{itemize}
\item The large integer module models large signed integer from 128 to 512 bits. The operations supported are bits operations , the four basic operators $(\times,+,/,-)$ , extended multiplications and bit shift operations. 
\item compile time polymorphism
\end{itemize}

\item \begin{itemize}
\item The polynomial module models polynomials up to 4 variables for dense and triangular polynomials, As previously, the polynomials usual  operations between polynomials and operations with monomials.  
\item polynomial can have large integer coefficient or other coefficient e.g. GMP
\end{itemize}

@@

\begin{itemize}
\item  We goal of this library is performed efficient arithmetic for polynomial with large coefficient and execute efficient inner product : multithread and/or GPU.
\item give an efficient library for the HP2C initiative code .e.g  SE
\end{itemize}
--------------------------------------------------------

@@

\section{Implementation of the library}
\subsubsection{memory layout}
\begin{itemize}
\item contrary to GMP  explain a static version of large arithmetic can be faster than a dynamic one
\item   VLI number are base on 64-bit integer array, size compile time polymorphism 
\item  polynomials constructed over boost::array, the vector of polynomial is std::vector
\item combined together we have a contiguous memory, and more precisely  SoA (show the figure). 
\item  It exists a second order name AoS, it does not allow a construction into module 
\end{itemize}

TRANSITION
\begin{itemize}
\item Polynomials operations necessitates large arithmetic operations for the coefficients ,
\item  coefficient operations must be fast, we need ASM 
\item this memory layout facilitates classical ASM 
\end{itemize}

\subsubsection{Kernels and assembly  language}

We applied algorithms based on childhood  experience for large arithmetic, for the specific range classical algorithms are enough.
The implementation necessitates a carry bit and an extended multiplication. The hardware supports
the carry bit in the status register, as the extended multiplications, but they are only accessible by the assembly language.
In assembly, these operations  are supported by the pair of mnemonics; \texttt{addq/adcq} for the addition; \texttt{subq/sbbq} for the subtraction, and \texttt{mul} for the extended multiplication.   
Combining these mnemonics, we reproduce easily long arithmetic algorithm, as described into the famous  ASM book of \cite{Hyde:2003:AAL:861534}.

@@

\begin{itemize}
\item  We does not use SIMD 
\item because the data order does not help
\item 
The disagreement  of this choice, the AoS order does not facilitate the usage of the SIMD instruction such as the AVX instructions of the Sandy Bridge x86-64 architecture.
Applying SIMD to the AoS, for an addition or a subtraction is a nonsense because there is no specific instruction to propagate carry bit between entries of the SIMD register.
It could be  apply  on multiplications for integer  larger than 1024-bit \cite{SIMD}, unfortunately, it does not correspond to our working range. Consequently, we do not 
use the multimedia instruction into the ASM kernels.
\end{itemize}

@@

\begin{itemize}
\item Introduce the capacity of the lib to perform inner product 
\item make a short math description 
\end{itemize}

>>>>

 THE LIB DOES
\begin{itemize}
\item The Very Large Integer library realizes large arithmetic , polynomials  and vector operation.  There are three 3 different modules interchangeable. 
\end{itemize}
Description of the 3 modules : vli, poly and vector
\begin{enumerate}
\item  \begin{itemize}
\item The large integer module models large signed integer from 128 to 512 bits. The operations supported are bits operations , the four basic operators $(\times,+,/,-)$ , extended multiplications and bit shift operations. 
\item compile time polymorphism
\end{itemize}
\item \begin{itemize}
\item The polynomial module models polynomials up to 4 variables for dense and triangular polynomials, As previously, the polynomials usual  operations between polynomials and operations with monomials.  
\item polynomial can have large integer coefficient or other coefficient e.g. GMP
\end{itemize}

@@

\begin{itemize}
\item  We goal of this library is performed efficient arithmetic for polynomial with large coefficient and execute efficient inner product : multithread and/or GPU.
\item give an efficient library for the HP2C initiative code .e.g  HTSE
\end{itemize}
--------------------------------------------------------

@@

\section{Implementation of the library}
\subsubsection{module VLI}
\begin{itemize}
\item contrary to GMP  explain a static version of large arithmetic can be faster than a dynamic one
\item   VLI number are base on 64-bit integer array, size compile time polymorphism 
\item VLI module performs arithmetic operations, it needs performable kernels
\item ASM Kernel

%\item  polynomials constructed over boost::array, the vector of polynomial is std::vector
%\item combined together we have a contiguous memory, and more precisely  SoA (show the figure). 
%\item  It exists a second order name AoS, it does not allow a construction into module

\end{itemize}

----------------------------------------------- \\
COPY/PAST from the LAST ASM\\
-----------------------------------------------\\
We applied algorithms based on childhood  experience for large arithmetic, for the specific range classical algorithms are enough.
The implementation necessitates a carry bit and an extended multiplication. The hardware supports
the carry bit in the status register, as the extended multiplications, but they are only accessible by the assembly language.
In assembly, these operations  are supported by the pair of mnemonics; \texttt{addq/adcq} for the addition; \texttt{subq/sbbq} for the subtraction, and \texttt{mul} for the extended multiplication.   
Combining these mnemonics, we reproduce easily long arithmetic algorithm, as described into the famous  ASM book of \cite{Hyde:2003:AAL:861534}.

@@

\begin{itemize}
\item  We does not use SIMD 
\item because the contiguous data does not help at all for CB propagation
\item 
%The disagreement  of this choice, the AoS order does not facilitate the usage of the SIMD instruction such as the AVX instructions of the Sandy Bridge x86-64 architecture.
%Applying SIMD to the AoS, for an addition or a subtraction is a nonsense because there is no specific instruction to propagate carry bit between entries of the SIMD register.
%It could be  apply  on multiplications for integer  larger than 1024-bit \cite{SIMD}, unfortunately, it does not correspond to our working range. Consequently, we do not 
%use the multimedia instruction into the ASM kernels.
\end{itemize}

\subsubsection{module polynomial}

short description of both type of polynomial

\begin{itemize}
\item construct on boost array
\item compile time polymorphism for dense, triangular
\item works whatever the coefficient
\item description of the implementation of the basic operations, the perf depends of coefficients implementation
\end{itemize}

\subsubsection{module vector}
\begin{itemize}
\item +- std::vector
\end{itemize}

\subsubsection{memory layout}

All elements modules combined together give the memory layout

\begin{itemize}
\item object conception imposes SoA
\item AoS is the good for SIMD ( global solution)
\item  with the flexibility of module we lost the SIMD 
\end{itemize}

@@

\begin{itemize}
\item Introduce the capacity of the lib to perform inner product  as external solver
\item make a short math description 
\end{itemize}
